---
title: "Data Consolidation Markdown-Clean"
author: "Evgeni Matveev, Elk River Alliance"
date: "2023-11-30"
output: html_document
---

```{r setup, include=FALSE}
install.packages("knitr") #need to install knitr first
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```


#STEP 1: Install Packages
```{r Install Packages}
#Remove any "#" preceeding the install.packages command and download packages onto your computer

#General R Packages
#might need to install Rtools
install.packages("rmarkdown")
install.packages("shiny")
install.packages("readxl")
install.packages("ggplot2")
install.packages("scales")
install.packages("dplyr")
install.packages("tidyverse")
install.packages("reshape2")
install.packages("ggpubr")
install.packages("patchwork")
install.packages("shiny")
install.packages("zoo")
install.packages("dplyr")
install.packages("ggmap")
install.packages("lubridate")
install.packages("ggrepel")
install.packages("plotly")

install.packages("sf")
install.packages("rgeos")
install.packages("sp")
install.packages("terra")


#Elevation Data
install.packages("elevatr")

#Water Survey Canada hydrometric data
install.packages("devtools")

install.packages("rems")

install.packages("tidyhydat")
library(devtools) #necessary for install_github()
install_github("bcgov/renmods",force=T)

library(pak)
pkg_install("bcgov/renmods")


#BC Data Catalogue
install.packages("bcdata")

#Weather Canada
install.packages("weathercan", 
                 repos = c("https://ropensci.r-universe.dev", 
                           "https://cloud.r-project.org"))

#For Columbia Basin Waterhub data
install.packages("ckanr")

```


#STEP 2: Load Packages
```{r Load Packages, message=FALSE, warning=FALSE}

library(rmarkdown)
library(sf)
#library(rgeos)
library(sp)
#library(terra)


library(ggplot2)
library(scales)
library(dplyr)
library(tidyverse)
library(reshape2)
library(ggpubr)
library(patchwork)
library(shiny)
library(zoo)
library(dplyr)
library(ggplot2)
library(ggmap)
library(lubridate)
library(readxl)
library(lubridate)
library(ggrepel)
library(plotly)

#Elevation data 
library(elevatr)

#Water Survey Canada
library(devtools)
library(rems)
library(tidyhydat)


#BC Data Catalogue
library(bcdata)

#Weather Canada
library(weathercan)

#Columbia Basin Waterhub Data
library(ckanr)

```


#STEP 3: Upload shape files of Area of Interest polygon (in this case, the Elk River Watershed) and stream shape file

```{r Part 1: Watershed Shape, fig.height=10}

#Make sure your .shp file and all related files (CPG,DBF, PRJ, SBN, SBX, and SHX) are in the same folder within your working directory
ERWshp<-st_read("./ERW Shape File/ElkRiverWatershed_CA_US.shp")

ERWshpPoly<-st_cast(ERWshp, "POLYGON")
st_zm(ERWshpPoly)

#Make sure your projection is set to NAD83 (crs code 4269)
st_transform(ERWshpPoly, crs = 4269)

```

```{r Part 2: Stream and Basin shapes, fig.height=10, results='hide'}

#Downloaded from:https://maps.gov.bc.ca/ess/hm/imap4m/

#Load Basin Polygons
Basin_polygon<-st_read("./Stream Network/NAMED_WATERSHED/FWA_NAMED_WATERSHEDS_POLY/FWNMDWTRSH_polygon.shp")
Basin_polygon<-st_cast(Basin_polygon, "POLYGON")
st_zm(Basin_polygon) #drop X and M dimensions
st_transform(Basin_polygon, crs = 4269)#Make sure your projection is set to NAD83 (crs code 4269)

ERWshpPoly2<-subset(Basin_polygon,GNIS_NAME=="Elk River")%>%select(geometry)

Basin_polygon_ERW<-Basin_polygon %>% 
  mutate(within_watershed = lengths(st_within(Basin_polygon,ERWshpPoly2)))%>%
  subset(within_watershed==1)
Basin_polygon_ERW<-Basin_polygon_ERW[order(-Basin_polygon_ERW$AREA_HA),] #Reorder by Basin Area (ha)


#Load stream Lines
Stream_lines<-st_read("./Stream Network/FWA_STREAM_NETWORKS_SP/FWSTRMNTWR_line.shp")
Stream_lines<-st_cast(Stream_lines, "LINESTRING")
st_zm(Stream_lines) #drop X and M dimensions
st_transform(Stream_lines, crs = 4269)#Make sure your projection is set to NAD83 (crs code 4269)
```

#STEP 4: Download Data

##A. Environmental Monitoring System
```{r Part 1-Current}

### PART 1 - CURRENT DATA (aka, last 2 years) ###

#Access EMS data from the last 2 years (THIS PART TAKES ABOUT 10-15 MINUTES)
CurrentEMS<-get_ems_data(which = "2yr", ask = FALSE)

########### if above data does not download 100%, retry 


#Collect coordinates and Site ID for all samples on EMS
AllEMSIDsCurrent<-CurrentEMS%>%
 filter(LOCATION_TYPE%in%c("RIVER,STREAM OR CREEK")& #this subsets the data to include only Rivers/Streams, delete if you'd like to access other location types
           SAMPLE_CLASS%in%c("Regular","BACKGROUND"))%>% #This subsets to only Regular and Background sample types (i.e., not replicates or blanks)
  distinct(EMS_ID,LONGITUDE,LATITUDE)%>% #Extract all the unique EMS sites and their coordinates
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4269) #Change the coordinates to be recognized as coordinates

#Categorize sites by whether they are in the shape polygon we defined in step 3, and select only those that are within the polygon
WithinERWCurrent <- AllEMSIDsCurrent %>% 
  mutate(within_watershed = lengths(st_within(AllEMSIDsCurrent, st_transform(ERWshpPoly, crs = 4269))))%>%
  subset(within_watershed==1)

#Create a list of all sites within your polygon
ERWSiteListCurrent<-as.character(WithinERWCurrent$EMS_ID)

#Subset your data to only include sites wihtin your polygon
CurrentERWdata<-CurrentEMS%>%
  filter(EMS_ID%in%ERWSiteListCurrent)

CurrentERWdata%>%subset(LOCATION_PURPOSE=="HYDROMETRIC")

```


```{r Part 2-Historic}
### PART 2 - HISTORIC DATA (aka, prior to two years ago) ###

#THIS SECTION TAKES A LONG TIME
download_historic_data() #When prompted to store a copy of ems data, type 1 or "Yes" into the console below
#if yes/no option doesn't appear, hit enter
######## if above data does not download 100%, retry
hist_db_con <- connect_historic_db()
hist_tbl <- attach_historic_data(hist_db_con)

#Download only the EMS id's and lat long first to filter out any sites outside of your polygon
filtered_historic<-hist_tbl%>%
 filter(LOCATION_TYPE%in%c("RIVER,STREAM OR CREEK")& #this subsets the data to include only Rivers/Streams, delete if you'd like to access other location types
           SAMPLE_CLASS%in%c("Regular","BACKGROUND"))%>% #This subsets to only Regular and Background sample types (i.e., not replicates or blanks)
  distinct(EMS_ID,LONGITUDE,LATITUDE)%>% #Extract all the unique EMS sites and their coordinates
  na.omit()%>%
  collect()

disconnect_historic_db(hist_db_con)

#Subset only sites within your polygon
AllEMSIDs<-st_as_sf(na.omit(filtered_historic),coords = c("LONGITUDE", "LATITUDE"), crs = 4269)
WithinERW <- AllEMSIDs %>% mutate(within_watershed = lengths(st_within(AllEMSIDs, st_transform(ERWshpPoly, crs = 4269))))%>%subset(within_watershed==1)

ERWSiteList<-as.character(WithinERW$EMS_ID)

#Now download the data only from sites that are within the watershed
hist_db_con <- connect_historic_db()
hist_tbl <- attach_historic_data(hist_db_con)

HistoricData<-hist_tbl%>%
  filter(EMS_ID%in%ERWSiteList)%>%
  collect() #THIS PART TAKES ABOUT 5 MINUTES

disconnect_historic_db(hist_db_con)


```

```{r Part 3-Combine}
### PART 3 - COMBINE CURRENT AND HISTORIC DATA###
EMSCurrentHistoricA <- bind_ems_data(CurrentERWdata, HistoricData)%>%distinct()

#Stream names have to be added manually due to inconsistency of naming convention
StreamNames<-read.csv("./EMS Monitoring Locations-edited.csv")%>%select(MONITORING_LOCATION, STREAM_NAME)  #upload excel spreadsheet
EMSCurrentHistoricB<-merge(EMSCurrentHistoricA, StreamNames,all.x=T, by="MONITORING_LOCATION")
EMSCurrentHistoricB$STREAM_NAME<-toupper(EMSCurrentHistoricB$STREAM_NAME)

#Add Data Owner (spreadsheet manually made by Anne-Caroline Kroeger)
EMSOwners<-read.csv("./EMS Data Owners.csv")%>%select(MONITORING_LOCATION_ID, PRESUMED_OWNER)%>%rename(EMS_ID=MONITORING_LOCATION_ID,DATA_OWNER=PRESUMED_OWNER)
EMSCurrentHistoricC<-merge(EMSCurrentHistoricB,EMSOwners,by="EMS_ID",all.x=T)


#Subset the most relevant data
EMS_complete<-EMSCurrentHistoricC %>% 
  subset(LOCATION_PURPOSE%in%c("TREND","BACKGROUND","HYDROMETRIC") &
           SAMPLE_CLASS=="Regular")%>%
  rename(
    SITE_ID = EMS_ID,
    DATE=COLLECTION_START,
    VALUE=RESULT,
    DETECTION_LIMIT=METHOD_DETECTION_LIMIT
    )%>%
  add_column(DATA_REPOSITORY="Environmental Monitoring System")%>%
  select(SITE_ID,MONITORING_LOCATION,STREAM_NAME,LONGITUDE,LATITUDE,DATE,DATA_REPOSITORY,PARAMETER,RESULT_LETTER,VALUE,UNIT,DATA_OWNER)


#THE EMS DATASET IS NOW COMPLETE

```

##B. Hydrometric Data

```{r Hydrometric Data Part 1-Water Survey Canada}
#Download data: THIS PART TAKES ABOUT 10-20 MINUTES
download_hydat() #when prompted, type "1" or "Yes" into your console to accept download

#Select only BC stations
BC_stns <- hy_stations() %>%
  filter(PROV_TERR_STATE_LOC == "BC")

#Set Lat Long as coordinates
BC_stns<-BC_stns%>%st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4269)

#Subset sites to be within the polygon
WaterSurveyERW_sites <- BC_stns %>% 
  mutate(within_watershed = lengths(st_within(BC_stns, st_transform(ERWshpPoly, crs = 4269))))%>%
  subset(within_watershed==1)

#Convert statWaterSurveyERWion name to stream name
WaterSurveyERW_sites$STREAM_NAME<- word(string = WaterSurveyERW_sites$STATION_NAME, start = 1, end = 2, sep = fixed(" "))

#Extract data for stations within the Elk River Watershed
WaterSurveyERW_data<-WaterSurveyERW_sites$STATION_NUMBER%>%hy_daily_flows()
WaterSurveyERW<-merge(WaterSurveyERW_data,WaterSurveyERW_sites,by="STATION_NUMBER")
WaterSurveyERW<-WaterSurveyERW%>%extract(geometry, c('LONGITUDE', 'LATITUDE'), '\\((.*), (.*)\\)', convert = TRUE) 

#add units based on Hydat r package documentation
WaterSurveyERW$UNIT<-"m^3/s"

#Converting Date to POSIXct format
WaterSurveyERW$Date<-as.POSIXct(WaterSurveyERW$Date,format="%Y-%m-%d")

#Add data owner (hydat only contains Water Survey Canada data)
WaterSurveyERW$DATA_OWNER<-"Water Survey Canada"

#Selecting Relevant Columns
WaterSurveyERW_complete<-WaterSurveyERW %>% 
  add_column(DATA_REPOSITORY="Water Survey Canada")%>%
  rename(
    SITE_ID = STATION_NUMBER,
    MONITORING_LOCATION = STATION_NAME,
    DATE=Date,
    PARAMETER=Parameter,
    VALUE=Value
    )%>%
  select(SITE_ID,MONITORING_LOCATION,STREAM_NAME,LONGITUDE,LATITUDE,DATE,DATA_REPOSITORY,PARAMETER,VALUE,UNIT,DATA_OWNER)
```


```{r Hydrometric Data Part 2-BC Data Catalogue}

#activate resource
bcdc_tidy_resources("9d9daae1-b379-4c92-b96d-6a81455caa74")

#The following comes from: https://catalogue.data.gov.bc.ca/dataset/hydrometric_data
#Each record has an associated resource number that can be seen in the URL of each subsection
#Discharge in this current chunk refers to volume of water in m3/s 
DischargeCurrent<- bcdc_get_data(record = '9d9daae1-b379-4c92-b96d-6a81455caa74', resource =
 'dd1fc2b7-22e3-46c5-848e-914b22586d5e') #NOT WORKING

Discharge20212022<-bcdc_get_data(record = '9d9daae1-b379-4c92-b96d-6a81455caa74', resource =
 '7c8239d1-e6f9-4c8b-9c14-c6ca8bd567c0')

Discharge20192021<-bcdc_get_data(record = '9d9daae1-b379-4c92-b96d-6a81455caa74', resource =
 '9ab0f5a3-3f1d-428c-86b2-9339163128c2')

Discharge20172019<-bcdc_get_data(record = '9d9daae1-b379-4c92-b96d-6a81455caa74', resource =
 '1273a53c-1617-472c-b549-fece71963c6e')

Discharge20152017<-bcdc_get_data(record = '9d9daae1-b379-4c92-b96d-6a81455caa74', resource =
 'ab59063b-8686-4804-a4b6-d876a398ce13')

DischargePre2015<-bcdc_get_data(record = '9d9daae1-b379-4c92-b96d-6a81455caa74', resource =
 '459cbc51-5003-4434-84f4-e4230e7ad3af')

#Combine all of the downloaded sources
BCDC_Discharge<-rbind(DischargeCurrent,Discharge20212022,Discharge20192021,Discharge20172019,Discharge20152017,DischargePre2015)

#Subset to polygBCDC_DISCHARGE_Coords<-BCDC_Discharge%>%select(`Location ID`,Latitude,Longitude)%>%distinct()%>%st_as_sf(coords = c("Longitude", "Latitude"), crs = 4269)

BCDC_DISCHARGE_ERW <- BCDC_DISCHARGE_Coords %>% mutate(within_watershed = lengths(st_within(BCDC_DISCHARGE_Coords, st_transform(ERWshpPoly, crs = 4269))))%>%subset(within_watershed==1)

BCDC_Discharge_ERW_DataA<-BCDC_Discharge%>%subset(`Location ID`%in%BCDC_DISCHARGE_ERW$`Location ID` & Value!="NaN")

#Change column names
BCDC_Discharge_ERW_Data<-BCDC_Discharge_ERW_DataA%>% 
  rename(
    SITE_ID = `Location ID`,
    MONITORING_LOCATION = `Location Name`,
    DATE=`Date/Time(UTC)`,
    PARAMETER=Parameter,
    VALUE=Value,
    UNIT=`Unit`,
    LATITUDE=Latitude,
    LONGITUDE=Longitude
    )

BCDC_Discharge_ERW_Data$STREAM_NAME<-BCDC_Discharge_ERW_Data$MONITORING_LOCATION%>%word(start = 1, end = 2, sep = fixed(" "))%>%as.character()%>%toupper()

#All Data Catalogue hydrometric stations are owned by Teck, based on research by Anne Caroline Kroeger
BCDC_Discharge_ERW_Data$DATA_OWNER<-"Teck"

#Finalise dataset
BCDC_Discharge_complete<-BCDC_Discharge_ERW_Data%>%
  add_column(DATA_REPOSITORY="BC Data Catalogue")%>%
  select(SITE_ID,MONITORING_LOCATION,STREAM_NAME,LONGITUDE,LATITUDE,DATE,DATA_REPOSITORY,PARAMETER,VALUE,UNIT,DATA_OWNER)

BCDC_Discharge_complete%>%select(SITE_ID,MONITORING_LOCATION)%>%unique()%>%write.csv("Teck BCDC IDs.csv")

```


##C. WaterHub Data
```{r Download Waterhub A}


ckanr_setup(url = "https://data.cbwaterhub.ca/")



#Load all potential packages
Packages_Table<-package_list_current(as="table",limit=500)%>%
  select(location,id,latitude,longitude,name,creator)%>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4269)

#Subset packages to Elk River Watershed
Packages_Table_ERW<-Packages_Table%>%
  mutate(within_watershed = lengths(st_within(Packages_Table, st_transform(ERWshpPoly, crs = 4269))))%>%
  subset(within_watershed==1)%>%
  as.data.frame()

#The following section loops through all the packages within the Elk River Watershed and packages and extracts only downloadable packages
x<-data.frame()
y<-data.frame()
z<-data.frame()

for (i in Packages_Table_ERW$id) {

DATA_OWNER<-Packages_Table_ERW%>%subset(id==i)%>%select(creator)%>%as.character()
STREAM_NAME<-Packages_Table_ERW%>%subset(id==i)%>%select(name)%>%word(start = 1, end = 2, sep = fixed("-"))%>%as.character()
MONITORING_LOCATION<-Packages_Table_ERW%>%subset(id==i)%>%select(location)%>%as.character()

x<-package_show(id=i, as="table")$resources

x<-as.data.frame(x)

if(is.null(x) || nrow(x) ==0) next

y<-x%>%subset(datastore_active=="TRUE")%>%
  select(id)%>%
  add_column("STREAM_NAME"=STREAM_NAME,"DATA_OWNER"=DATA_OWNER,"MONITORING_LOCATION"=MONITORING_LOCATION)

z<-rbind(z,y)

}

#Modify stream names
z$STREAM_NAME<-gsub("-"," ",z$STREAM_NAME)%>%toupper()

#The following section loops through all packages and extracts the data
#The data needs to be consolidated into a single source
#Because the different data tables have different structures, they need to be manually unified

ValueNames<-c("values","water_temperature","average_water_temperature","average_water_temp","flow", "discharge", "water_level") 
UnitNames<-c("flow_units","unit_of_measurement","average_water_temperature_units","water_temperature_units","average_water_temp_units", "metres", "water_level_units", "discharge_units")
SiteIDNames<-c("site_id","SITE_ID","site_ID")
LongitudeNames<-c("longitude","Longitude","site_longitude")
LatitudeNames<-c("latitude","Latitude","site_latitude")
DateNames<-c("date_time_sampled","date_time","")

analyte <- "other"

ERApackagelist<-list()
WaterhubData_complete_unformatted<-data.frame()

for (j in z$id) {

name<-z%>%subset(id==j)%>%select(STREAM_NAME)%>%as.character()
owner<-z%>%subset(id==j)%>%select(DATA_OWNER)%>%as.character()
location<-z%>%subset(id==j)%>%select(MONITORING_LOCATION)%>%as.character()

table<-ckan_fetch(x=paste0("https://data.cbwaterhub.ca/dataset/80804406-bce2-417b-a62a-ff0981cec6a1/resource/",j,"/download/era_temp_liz001_20210707_20220127.csv"))


if(j=="91e6da27-7bb5-4e3b-a12e-6fc5c56fbe4d"){
  table<-table%>%mutate(water_temperature_units="degrees Celsius",latitude="49.501079",longitude="-115.174318")
}


tableB<- table%>%
  mutate(DATA_OWNER=owner,STREAM_NAME=name,MONITORING_LOCATION=location)%>%
  rename("VALUE"=any_of(ValueNames),
         "SITE_ID"=any_of(SiteIDNames),
         "UNIT"=any_of(UnitNames),
         "LATITUDE"=any_of(LatitudeNames),
         "LONGITUDE"=any_of(LongitudeNames),
         "DETECTION_LIMIT"=any_of("detection_limit"),
         "DATE"=any_of(DateNames))%>%
  dplyr::rename("VALUE"=any_of("VALUE1"),
         "UNIT"=any_of("UNIT2"))

tableB <- tableB %>% filter()



tableC<-as.data.frame(tableB)%>%
  mutate(PARAMETER=ifelse(UNIT%in%c("degrees Celsius","degrees Celcius"),"Temperature",ifelse(UNIT%in%c("meter3 per second","metres3 per second"),"Discharge",analyte)))%>%
  select(DATE,SITE_ID,MONITORING_LOCATION,LATITUDE,LONGITUDE,VALUE,UNIT,DATA_OWNER,STREAM_NAME,PARAMETER)

WaterhubData_complete_unformatted<-rbind(tableC,WaterhubData_complete_unformatted)
  
}

#Remove the letter T from dates
WaterhubData_complete_unformatted$DATE<-sub("T"," ",WaterhubData_complete_unformatted$DATE)
```


Some dates have been entered as yy-dd-mm instead of yyyy-dd-mm. Excel will autocorrect dates to correct format, but Google Sheets will not if. IF USING SHEETS, run following section to correct date formatting.
```{r}
WaterhubData_complete <-  WaterhubData_complete %>% mutate(YEAR = str_extract(DATE, "^[^ ]+"))
WaterhubData_complete <-  WaterhubData_complete %>% mutate(TIME = str_remove(DATE, "^[^ ]+ "))


i=1
j=1
for (i in 1:nrow(WaterhubData_complete)){
     
  if (nchar(WaterhubData_complete$YEAR[j]) < 10){
    WaterhubData_complete$YEAR[j] <- paste(20, WaterhubData_complete$YEAR[j], sep="")}
  j=j+1
}


WaterhubData_complete$DATE <- NULL

i=1
j=1
for (i in 1:nrow(WaterhubData_complete)){
     
    WaterhubData_complete$DATE[j] <- paste(WaterhubData_complete$YEAR[j], WaterhubData_complete$TIME[j] , sep=" ")
  j=j+1
}

WaterhubData_complete$YEAR <- NULL
WaterhubData_complete$TIME <- NULL

WaterhubData_complete2 <- WaterhubData_complete
```



```{r}
write.csv(WaterhubData_complete,"WaterhubData_complete_unformatted.csv")

```

#To reformat the date column, the dataset needs to be saved as a CSV, opened in Excel, and the date needs to be reformatted there
#Open the csv in Excel, select the DATE column, go to custom formatting, and format it as yyyy-dd-mm hh:mm:ss
#Then Save As the reformatted sheet as WaterhubData_complete_formatted.csv" and go to the next chunk
write.csv(WaterhubData_complete_unformatted,"WaterhubData_complete_unformatted.csv")




```{r WaterHub B}

#Reupload the spreadsheet with the reformatted date
WaterhubData_complete<-read.csv("C:/Users/user/Desktop/ERA/Partner data review/MonCol Code and Data/Data Consolidation Markdown/WaterhubData_complete_formatted.csv")

```

Ensure structure of data is correct
```{r}
WaterhubData_complete$DATE<-as.POSIXct(WaterhubData_complete$DATE,format="%Y-%d-%m %H:%M:%S")
WaterhubData_complete$VALUE<-as.numeric(WaterhubData_complete$VALUE)
WaterhubData_complete$RESULT_LETTER<-NA
WaterhubData_complete$DATA_REPOSITORY<-"Columbia Basin Water Hub"

```

##D. Weather Canada Data

##E. CABIN data
```{r CABIN Data: Site Info}

options(timeout=6000) #set timeout time to 10 mins, otherwise the download won't complete

#Download site information for Pacific drainage basin 
CABINSiteData<-read.csv("https://cabin-rcba.ec.gc.ca/Cabin/opendata/cabin_study_data_mda08_1987-present.csv",fileEncoding = "UTF-16")

CABINSiteData%>%subset(Site.Site=="ELB01A")

#Subset site data to watershed polygon
CABIN_coords<-CABINSiteData%>%select(`Site.Site`,Latitude,Longitude,SiteName.Nomdusite,SiteVisitID.IdentifiantdeVisite)%>%
  distinct()%>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4269)

CABIN_coords_ERW_A <- CABIN_coords %>% 
  mutate(within_watershed = lengths(st_within(CABIN_coords, st_transform(ERWshpPoly, crs = 4269))))%>%
  subset(within_watershed==1)

#Some SiteIDs are shared between multiple unrelated sites. 
##The code below calculates how many unique location descriptions (SiteName.Nomdusite) are associated with each SiteID (Site.Site)
##Only completely unique site locations can be used to subset Benthic and Habitat data
#NOTE: Elbow River has to be remove because the coordinates were not entered correctly, putting it into the Elk River Watershed
#NOTE: The same site names have many different coordinates 
CABIN_coords_ERW_C<-CABINSiteData%>%
  add_column(DATA_REPOSITORY="CABIN")%>%
  subset(SiteVisitID.IdentifiantdeVisite%in%CABIN_coords_ERW_A$SiteVisitID.IdentifiantdeVisite&SiteName.Nomdusite!="Elbow River")%>%
  rename(LATITUDE=Latitude,LONGITUDE=Longitude,MONITORING_LOCATION=SiteName.Nomdusite,SITE_ID=Site.Site)%>%
  select(LATITUDE, LONGITUDE, MONITORING_LOCATION,SiteVisitID.IdentifiantdeVisite,SITE_ID)



CABIN_coords_ERW_C$MONITORING_LOCATION<-toupper(CABIN_coords_ERW_C$MONITORING_LOCATION)

```

```{r CABIN data: Benthics}

#Download Benthic Invertebrate data for Pacific drainage basin (TAKES A LONG TIME 5-10 mins)
CABINBenthicData<-read.csv("https://cabin-rcba.ec.gc.ca/Cabin/opendata/cabin_benthic_data_mda08_1987-present.csv",fileEncoding = "UTF-16")


#Subset data to only include Elk River Watershed sites
CABINBenthicData_ERW<-subset(CABINBenthicData,SiteVisitID.IdentifiantdeVisite%in%CABIN_coords_ERW_C$SiteVisitID.IdentifiantdeVisite)
CABINBenthicData_ERW_B<-merge(CABINBenthicData_ERW,CABIN_coords_ERW_C,by="SiteVisitID.IdentifiantdeVisite")


#Subset to 3 minutes of kick time and 400nm mesh size
#Rename columns, then summarize counts of benthics by Order
CABINBenthicData_complete<-CABINBenthicData_ERW_B%>%
  mutate("DATE"=as.Date(JulianDay.JourJulien,origin=paste0(Year.Année,"-01-01"))) %>% #Note, this date is an estimate, but accurate to within a day or two
  subset(MeshSize.Maillage==400&KickTime.Périodedelapassedufilettroubleau==3)%>%
  rename("ORDER"=Order.Ordre,
         "VALUE"=Count.Décompte)%>%
  add_column("PARAMETER"="Benthic Invertebrates",
             "UNIT"="Count",
             DATA_REPOSITORY="CABIN")%>%
  group_by(Year.Année,SITE_ID,ORDER) %>%
  summarise(VALUE=sum(VALUE),Year.Année,SITE_ID,ORDER,DATE,DATA_REPOSITORY,PARAMETER,UNIT,MONITORING_LOCATION,LATITUDE,LONGITUDE) %>%
  distinct()%>%
  select(DATE,SITE_ID,MONITORING_LOCATION,DATA_REPOSITORY,LATITUDE,LONGITUDE,PARAMETER,UNIT,ORDER,VALUE)

#Reorder main orders so that EPT is first
FillOrder<-c("Ephemeroptera","Plecoptera", "Trichoptera","Anthoathecatae", "Coleoptera", "Collembola", "Diptera", "Lumbriculida", "Neotaenioglossa", "Sarcoptiformes", "Trombidiformes", "Tubificida")  

CABINBenthicData_complete$ORDER<-factor(CABINBenthicData_complete$ORDER,levels=FillOrder)


#BENTHIC DATA NOW COMPLETE  

```

```{r CABIN Data: Habitat}

#Download water quality data (habitat data) for Pacific drainage basin (TAKES A LONG TIME: 10-15 mins)
CABINHabitatData<-read.csv("https://cabin-rcba.ec.gc.ca/Cabin/opendata/cabin_habitat_data_mda08_1987-present.csv",fileEncoding = "UTF-16")


CABINHabitatData_ERW<-subset(CABINHabitatData,SiteVisitID.IdentifiantdeVisite%in%CABIN_coords_ERW_C$SiteVisitID.IdentifiantdeVisite)

CABINHabitatData_ERW_B<-merge(CABINHabitatData,CABIN_coords_ERW_C,by="SiteVisitID.IdentifiantdeVisite")

CABINHabitatData_complete<-CABINHabitatData_ERW_B%>%
  subset(Type.Type=="Water Chemistry")%>% #This selects only water chemistry data (i.e., heavy metals, nutrients, TSS, pH)
  mutate("DATE"=as.Date(JulianDay.JourJulian,origin=paste0(Year.Année,"-01-01")))%>%
  add_column(DATA_REPOSITORY="CABIN")%>%
  rename(PARAMETER=VariableDescription,
             UNIT=Unit.Unité,
             VALUE=Value.Valeur,
             DETECTION_LIMIT=MDL)%>%
  select(DATE,SITE_ID,MONITORING_LOCATION,LATITUDE,DATA_REPOSITORY,LONGITUDE,PARAMETER,UNIT,VALUE,DETECTION_LIMIT)

#HABITAT DATA NOW COMPLETE
CABINHabitatData_complete$PARAMETER%>%unique() ############## reason to print?




```



#STEP 5: Consolidate Data
```{r Data Consolidation}
all_data<-bind_rows(EMS_complete,
                    WaterSurveyERW_complete,
                    BCDC_Discharge_complete,
                    WaterhubData_complete,
                    CABINBenthicData_complete,
                    CABINHabitatData_complete)

#Add Day Month and Year as numeric columns
all_data$day<- format(all_data$DATE, "%j")%>%as.numeric()
all_data$month<- format(all_data$DATE, "%m")%>%as.numeric()
all_data$year<- format(all_data$DATE, "%Y")%>%as.numeric()

Data_by_owner<-all_data%>%select(SITE_ID,DATA_OWNER)%>%unique()%>%group_by(DATA_OWNER)%>%summarise(SiteNumber=n())


########### should you clarify what these lines below are for?
all_data$PARAMETER%>%unique()%>%length()

all_data$DATE%>%na.omit()%>%max()

```


#STEP 6: Add site elevation Data
```{r Site Elevation}
#THIS TAKES ABOUT 5-10 MINUTES#

#The same site IDs may have slightly different coordinates, which is important for CABIN but not as important for elevation data
#Here we estimate the average location of each site ID
SiteCoordinates<-all_data%>%
  select(SITE_ID,LATITUDE,LONGITUDE)%>%
  unique()%>%
  group_by(SITE_ID)%>%
  mutate(LATITUDE=mean(LATITUDE),LONGITUDE=mean(LONGITUDE))%>%
  unique()%>%
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4269)

#Adding Site Elevations
SiteElevations <- get_elev_point(SiteCoordinates, prj = 4269, src = "aws",z=13)

data.frame(SiteElevations)

all_data_w_elevations<-merge(all_data, SiteElevations[ , c("SITE_ID", "elevation")], by="SITE_ID")

all_data<-all_data_w_elevations

```

#STEP 7: Add basin data
```{r Add Basin Data}

Site_Locations<-all_data%>%
          select(SITE_ID,LONGITUDE,LATITUDE)%>%
          distinct()%>%
          st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4269)

#Determine all named basins a site is within
Station_in_Basins<-apply(st_intersects(Basin_polygon_ERW, Site_Locations, sparse = FALSE), 2, 
                   function(col) { 
                   Basin_polygon_ERW[which(col), ]$GNIS_NAME
               })

#Create dataframe of the main basin for each site
MainBasin<-data.frame()
for (i in c(1:length(Site_Locations$SITE_ID))) {
  
if (length(Station_in_Basins[[i]])==1){
  x<-paste(Station_in_Basins[[i]], collapse = ', ')
  }
else {
  x<-paste(Station_in_Basins[[i]][2], collapse = ', ')
}
MainBasin<-rbind(MainBasin,x)
}

PtsWBasins<-cbind(Site_Locations,"GNIS_NAME"=MainBasin[,1])

all_data_w_basins<-merge(all_data, PtsWBasins[ , c("SITE_ID", "GNIS_NAME")], by="SITE_ID")

all_data<-all_data_w_basins

```


#STEP 8: Write CSVs
```{r Write CSV}

paste0("Consolidated Data:",Sys.Date(),".csv")




select_parameters<-read.csv("All Parameters_SELECTED.csv")

included_parameters<-select_parameters%>%
  subset(INCLUDE==2)%>%
  select(PARAMETER,PARAMETER_GENERAL)%>%
  unique()

all_data_selected<-all_data%>%
  select(-geometry.x,-geometry.y,-Year.Année,-X)%>%
  subset(PARAMETER%in%included_parameters$PARAMETER)%>%
  merge(.,included_parameters,by="PARAMETER")


all_data_selected$PARAMETER_GENERAL.y <- NULL

all_data_tiny<-all_data_selected%>%
  subset(PARAMETER_GENERAL%in%c("Nitrate(NO3) + Nitrite(NO2) Dissolved",
                                "Calcium Total",
                                "Carbon Organic Total",
                                "Benthic Invertebrates",
                                "Selenium Total",
                                "Turbidity",
                                "Flow"))%>%unique()

write.csv(all_data,"./Consolidated Data-Full.csv")
write.csv(all_data_selected,"Consolidated Data-Selected.csv")
write.csv(all_data_tiny,"Consolidated Data-Tiny.csv")


```
